Queue Aware Precoder Design for Space Frequency Resource Allocation

### 2014-06-26 Report/joint_wsrm_formulation.tex
ver_0.0.1

The \ac{WSRM} technique is predominantly used to design the precoders for the \ac{MIMO} \ac{BC} channels to maximize the weighted sum throughput in the downlink direction. The weights are used to provide different priorities to the users based on their requirements and \ac{QoS} criterion. Since the objective of this work is mainly concerned with the reduction of the number of queued packets, we use the length of the queued packets for each user as the respective weights. Using the length of the queued packets as the corresponding weights is obtained from the well known algorithm, called the \textit{backpreassure algorithm}, which is used to decide the routing decisions for each node in the network to efficiently send the packets to the desired destination with minimal packet drops and delay. The backpreassure algorithm is the outcome of minimizing the Lyapunov drift conditioned over the current queue states \me{\mbf{Q}(i - 1)}, where \me{\mbf{Q}} is the vector formed by stacking the number of queued packets of each user in the system. Assuming the queue length grows according to \eqref{eqn-2a} for all the users, the Lyapunov drift is given by \cite{georgiadis2006resource,neely2010stochastic}
\begin{equation}
L(i) = \frac{1}{2} \, \| \mbf{Q}(i) \|^2_2 = \frac{1}{2} \, \sum_{k \in \mc{U}} Q_k(i)^2.
\end{equation}
The conditional Lyapunov drift is given by taking expectation over all channel realizations as
\begin{equation}
\Delta(i) = \mathrm{E}_H \left [ L(i) - L(i - 1) | \mbf{Q}(i - 1) \right ],
\label{eqn-3.1.0}
\end{equation}
since the transmission depends on the channel condition. Now, by substituting \eqref{eqn-2a} in \eqref{eqn-3.1.0}, the resulting expression can be upper bounded by
\begin{equation}
\Delta(i)\leq \sum_{k \in \mc{U}} \, \mathrm{E}_H \left [ \mbf{Q}(i-1) \left ( \mbf{\lambda}(i) - \mbf{t}(i) \right ) | \mbf{Q}(i-1) \right ] + \mathrm{E}_H \left [ \mbf{\lambda}^2(i-1) \right ] + \mathrm{E} \left [ \mbf{t}^2(i-1) \right ],
\end{equation}
where \me{\mbf{\lambda}} and \me{\mbf{t}} are the stacked arrivals and transmissions of all users. Since the expectation is conditioned over \me{\mbf{Q}(i-1)} and the second order statistics of the arrivals and the transmissions are bounded \cite{neely2010stochastic}, the drift minimization solution is given as
\begin{equation}
\underset{\mbf{t}(i)}{\text{minimize}} \sum_{k \in \mc{U}} \, Q_k(i-1) \: t_k(i).
\label{backpreassure}
\end{equation}

>>>>

The \ac{WSRM} technique is predominantly used to design the precoders for the \ac{MIMO} \ac{BC} channels to maximize the weighted sum throughput in the downlink direction. The weights are used to provide different priorities to the users based on their requirements and \ac{QoS} criterion. Since the objective of this work is mainly concerned with the reduction of the number of queued packets, we use the length of the queued packets for each user as the respective weights. Using the length of the queued packets as the corresponding weights is obtained from the well known algorithm, called the \textit{backpreassure algorithm}, which is used to decide the routing decisions for each node in the network to efficiently send the packets to the desired destination with minimal packet drops and delay. The backpreassure algorithm is the outcome of minimizing the Lyapunov drift conditioned over the current queue states \me{\mbf{Q}(i - 1)}, where \me{\mbf{Q}} is the vector formed by stacking the number of queued packets of each user in the system. Assuming the queue length grows according to \eqref{eqn-2a} for all the users, the Lyapunov drift conditioned over different channel realizations is given by \cite{georgiadis2006resource,neely2010stochastic}
\begin{equation} \label{eqn-3.1.0}
\Delta(i) = \frac{1}{2} \, \sum_{k \in \mc{U}} \, \mathrm{E}_H \Big{[}  \left ( Q_k(i)^2 - Q_k(i-1)^2 \right ) \vert \, \mbf{Q}(i - 1) \Big{]}.
\end{equation}
Now, by substituting \eqref{eqn-2a} in \eqref{eqn-3.1.0}, the resulting expression can be upper bounded by
\begin{equation}
\Delta(i)\leq \mathrm{E}_H \Big{[} \mbf{Q}(i-1) \left \lbrace \mbfa{\lambda}(i) - \mbf{t}(i) \right \rbrace | \mbf{Q}(i-1) \Big{]} + \mathrm{E}_H \left [ \mbfa{\lambda}^2(i-1) \right ] + \mathrm{E}_H \left [ \mbf{t}^2(i-1) \right ],
\end{equation}
where \me{\mbfa{\lambda}} and \me{\mbf{t}} are the stacked arrivals and transmissions of all users. Since the expectation is conditioned over \me{\mbf{Q}(i-1)} and the second order statistics of the arrivals and the transmissions are bounded \cite{neely2010stochastic}, the drift minimization solution is given as
\begin{equation}
\underset{\mbf{t}(i)}{\text{minimize}} \sum_{k \in \mc{U}} \, Q_k(i-1) \: t_k(i).
\label{backpreassure}
\end{equation}

### 2014-06-26 Report/problem_formulation.tex
ver_0.0.1

The precoder design problem can be given minimizing the absolute value of the deviation in \eqref{eqn-4.2} raised to the exponent \me{q} as
\me{\text{minimize} \: \sum_{k \in \mc{U}} \, | v_k |^q}. The exponent \me{q} plays a vital role in the final allocation, which will be discussed later. Now, the problem of weighted queued packet minimization formulated as a $q$-norm minimization, is given by
\begin{IEEEeqnarray}{rCl}\label{eqn-3}
\underset{\substack{\mvec{m}{l,k,n},\\\mvec{w}{l,k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}  \|_q\IEEEyessubnumber \\
\text{subject to} & \quad&\sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b, \IEEEyessubnumber \label{eqn-4.3}
\end{IEEEeqnarray}
where \me{\tilde{v}_k \triangleq a_k^{1/{q}}v_k}, \me{a_k} is the weighting factor which is incorporated to control user priority based on their respective \ac{QoS}, \me{\gamma_{l,k,n}} is defined in \eqref{eq:SINR}, and \me{\mbf{M}_{k,n} \triangleq [ \, \mvec{m}{1,k,n} \, \mvec{m}{2,k,n} \dotsc \mvec{m}{L,k,n} \,]} comprises the beamformers associated with the user \me{k} for \me{n^\mathrm{th}} sub-channel transmission. It can be easily extended for user specific streams \me{L_k} instead of using the common \me{L} streams for all users. The expression for \me{t_{l,k,n}} is due to the assumption of Gaussian signaling, so that the maximum achievable rate is \me{\log_2(1 + \gamma_{l,k,n})} for a given \ac{SINR} \me{\gamma_{l,k,n}}. In \eqref{eqn-4.3}, we consider the sum power constraint for each \ac{BS} across all sub-channels. The proposed solution presented in this section also applies to the sub-channel power constraint by replacing \eqref{eqn-4.3} by corresponding formulations. Before proceeding further, we note that the constraint in \eqref{eqn-4.2} is handled implicitly by the definition of the \me{q}-norm in the objective of \eqref{eqn-3}. As a proof, suppose that \me{t_k>Q_k} for a certain \me{k} at optimum, i.e., \me{-{v}_k=t_k-Q_k>0}. Then there exists \me{\delta_k>0} such that \me{-{v}^{\prime}_k=t^{\prime}_k-Q_k<-{v}_k} where \me{t^{\prime}_k=t_k-\delta_k}. Since \me{\|\tilde{\mbf{v}}\|_q=\| |\tilde{\mbf{v}}| \|_q=\||-\tilde{\mbf{v}}|\|_q}, this means that the newly created vector \me{\mbf{t}^{\prime}} achieves a smaller objective which contradicts with the fact that an optimal solution has been obtained. We comment on the choice of  the norm \me{q} on the objective as below \cite{berry2004cross,qps_cioffi}.
\begin{itemize}
\item With \me{q = 1}, the objective results in greedy allocation \textit{i.e}, emptying the queue of users with good channel condition before considering the users with worse channel conditions. As a special case, it is easy to see that \eqref{eqn-3} reduces to the \ac{Q-WSRM} problem \eqref{gen_sum} when the queue size is large enough for all users.
\item With \me{q = 2}, the objective prioritizes users with higher queued packets before considering the users with a smaller number of backlogged packets. This is ideal for the delay limited scenario when the packet arrival rates of the users are similar, since the backlogged packets is proportional to the delay in the transmission following the Little's law \cite{neely2010stochastic}.
\item With \(q = \infty \), the objective minimizes the maximum number of queued packets among users with the current transmission, thereby providing queue fairness by allocating the resources proportional to the number of backlogged packets.
\end{itemize}

>>>>

The precoder design problem can be given minimizing the absolute value of the deviation in \eqref{eqn-4.2} raised to the exponent \me{q} as
\me{\text{minimize} \: \sum_{k \in \mc{U}} \, | v_k |^q}. The exponent \me{q} plays a vital role in the final allocation, which will be discussed later. Now, the problem of weighted queued packet minimization formulated as a $q$-norm minimization, is given by
\begin{IEEEeqnarray}{rCl}\label{eqn-3}
\underset{\substack{\mvec{m}{l,k,n},\\\mvec{w}{l,k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}  \|_q\IEEEyessubnumber \\
\text{subject to} & \quad&\sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b, \IEEEyessubnumber \label{eqn-4.3}
\end{IEEEeqnarray}
where \me{\tilde{v}_k \triangleq a_k^{1/{q}}v_k}, \me{a_k} is the weighting factor which is incorporated to control user priority based on their respective \ac{QoS}, \me{\gamma_{l,k,n}} is defined in \eqref{eq:SINR}, and \me{\mbf{M}_{k,n} \triangleq [ \, \mvec{m}{1,k,n} \, \mvec{m}{2,k,n} \dotsc \mvec{m}{L,k,n} \,]} comprises the beamformers associated with the user \me{k} for \me{n^\mathrm{th}} sub-channel transmission. It can be easily extended for user specific streams \me{L_k} instead of using the common \me{L} streams for all users. The expression for \me{t_{l,k,n}} is due to the assumption of Gaussian signaling, so that the maximum achievable rate is \me{\log_2(1 + \gamma_{l,k,n})} for a given \ac{SINR} \me{\gamma_{l,k,n}}. In \eqref{eqn-4.3}, we consider the sum power constraint for each \ac{BS} across all sub-channels.

### 2014-06-26 Report/wsrm_formulation.tex
ver_0.0.1

>>>>

The proposed solution presented in this section also applies to the sub-channel power constraint by replacing \eqref{eqn-4.3} by corresponding formulations. Before proceeding further, we note that the constraint in \eqref{eqn-4.2} is handled implicitly by the definition of the \me{q}-norm in the objective of \eqref{eqn-3}. As a proof, suppose that \me{t_k>Q_k} for a certain \me{k} at optimum, i.e., \me{-{v}_k=t_k-Q_k>0}. Then there exists \me{\delta_k>0} such that \me{-{v}^{\prime}_k=t^{\prime}_k-Q_k<-{v}_k} where \me{t^{\prime}_k=t_k-\delta_k}. Since \me{\|\tilde{\mbf{v}}\|_q=\| |\tilde{\mbf{v}}| \|_q=\||-\tilde{\mbf{v}}|\|_q}, this means that the newly created vector \me{\mbf{t}^{\prime}} achieves a smaller objective which contradicts with the fact that an optimal solution has been obtained. We comment on the choice of  the norm \me{q} on the objective as below \cite{berry2004cross,qps_cioffi}.
\begin{itemize}
\item With \me{q = 1}, the objective results in greedy allocation \textit{i.e}, emptying the queue of users with good channel condition before considering the users with worse channel conditions. As a special case, it is easy to see that \eqref{eqn-3} reduces to the \ac{Q-WSRM} problem \eqref{gen_sum} when the queue size is large enough for all users.
\item With \me{q = 2}, the objective prioritizes users with higher queued packets before considering the users with a smaller number of backlogged packets. This is ideal for the delay limited scenario when the packet arrival rates of the users are similar, since the backlogged packets is proportional to the delay in the transmission following the Little's law \cite{neely2010stochastic}.
\item With \(q = \infty \), the objective minimizes the maximum number of queued packets among users with the current transmission, thereby providing queue fairness by allocating the resources proportional to the number of backlogged packets.
\end{itemize}
